A Switching Planner for Combined Task and Observation Planning

### 2010-11-25 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/conc.tex
tiny changes.

One theoretical criticism of switching continual planning concerns
interleaved sequential and decision-theoretic sessions failing to make
any progress towards the POMDP objectives. For example, in the worst
case we could have each sequential session producing an identical
plan, and each decision-theoretic session rejecting it without further
sensing. Although this has not been an issue in our work so far, it
must be dealt with rigorously moving forward. We propose a good way of
mitigation this problem is by using a {\em motivational} component
that maintains a {\em dynamic} reward model for sensing.

@@

In this paper we present an approach to continual planing that uses
two planning systems. The first

>>>>

One theoretical criticism of switching continual planning concerns
interleaved sequential and decision-theoretic sessions failing to make
any progress towards the POMDP objectives. For example, in the worst
case we could have each sequential session producing an identical
plan, and each decision-theoretic session rejecting it without further
sensing. Although this has not been an issue in our work so far, it
must be dealt with rigorously moving forward. We suggest that a good
way of mitigate this problem is by using a {\em motivational}
component that maintains a {\em dynamic} reward model for sensing.

@@

In this paper we present an approach to continual planning that uses
two planning systems. The first

### 2010-11-25 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex
tiny changes.

We model a process state $\state$ as the set of propositions that are
true of the state. Notationally, we have $\state \subseteq
\props$. State change is induced by application of actions. A
stochastic action $\stochAction \in \stochActions$ is applicable if
its precondition $\poss(\stochAction)$, a set of propositions, are
satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
\state$. We write $\stochActions(\state)$ for set of actions
applicable at state $\state$.  If action $\stochAction \in
\stochActions(\state)$ is applied at \state, nature chooses one
element amongst a small set of deterministic outcomes
$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
nature takes outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
effect on the state given in terms of two lists of propositions. The
add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
$\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
..,\prop_m]$ is chosen by nature when \stochAction\ is
applied at state $\state$, then the resultant state is $ ( \state \cup
\add(\detAction) ) \backslash \delete(\detAction)$ -- I.e.,
propositions from $\add(\detAction)$ are added to $\state$, and those
from $\delete(\detAction)$ are removed from $\state$.

@@

Finally, we make a number of fairly standard assumptions. First, that
action execution and sensing occurs instantaneously, and that only one
action can be applied at a plan-step. Second, it can arise that in
some propositional states an action is applicable, and in others it is
not. Moreover, a belief-state can assign non-zero probability to
states where that applicability holds, and states where it does
not. We differ
from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
because we forbid the execution of actions that are not applicable in
any state of the current belief.  Also,
unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
we do incorporate a PPDDL-like default semantics, supposing that when
an action \stochAction\ is executed at $\state_i$, that it has no
effect on states $\state_i$ if $\poss(\stochAction) \not\subseteq
\state_i$.

@@

An optimal solution to a finite-horizon POMDP is a contingent plan,
and can be expressed as a mapping from observation histories to
actions. Although suboptimal in general, useful plans can also take a
classical sequential format. This is the case in {\em conformant}
planning, where the objective is to find a sequence of actions that
achieves a goal ---I.e., reaches a state that satisfies some given
Boolean condition--- with probability $1$.  More generally, a useful
formalism for representing solutions to POMDPs is the finite-state
controller (FSC). This is a three-tuple $\langle \nodes, \selection,
\transition \rangle$ where: $\node \in \nodes$ is a set of controller
states, $\selection_\node(\action) = P(\action | \node)$ gives the
probability that the controller prescribes \action\ when in state
\node, and $\transition_\node(\action, \observation, \node') =
P(\node'|\node, \action, \observation)$ gives the probability the
controller transitions to state $\node'$, supposing we execute
\action\ in state \node\ receiving observation \observation. When
acting for finite $N$ steps from controller state $n$, the value of a
controller corresponds to the expected cumulative reward:

>>>>

We model a process state $\state$ as the set of propositions that are
true of the state. Notationally, we have $\state \subseteq
\props$. State change is induced by application of actions. A
stochastic action $\stochAction \in \stochActions$ is applicable if
its precondition $\poss(\stochAction)$, a set of propositions, are
satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
\state$. We write $\stochActions(\state)$ for set of actions
applicable at state $\state$.  If action $\stochAction \in
\stochActions(\state)$ is applied at \state, nature chooses one
element amongst a small set of deterministic outcomes
$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
nature takes outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
effect on the state given in terms of two lists of propositions. The
add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
$\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
..,\prop_n]$ and $\delete(\detAction) := [\prop_{n+1},
..,\prop_m]$ is chosen by nature when \stochAction\ is
applied at state $\state$, then the resultant state is $ ( \state \cup
\add(\detAction) ) \backslash \delete(\detAction)$ -- i.e.,
propositions from $\add(\detAction)$ are added to $\state$, and those
from $\delete(\detAction)$ are removed from $\state$.

@@

Finally, we make a number of fairly standard assumptions. First, that
action execution and sensing occurs instantaneously, and that only one
action can be applied at a plan-step. Second, it can arise that in
some propositional states an action is applicable, and in others it is
not. Moreover, a belief-state can assign non-zero probability to
states where that applicability holds, and states where it does
not. We differ
from~\citeauthor{younes:littman:04}~(\citeyear{younes:littman:04}),
because we forbid the execution of actions that are not applicable in
any state of the current belief.  Also,
unlike~\citeauthor{hoffmann:brafman:2006}~(\citeyear{hoffmann:brafman:2006}),
we do incorporate a PPDDL-like default semantics, treating an action
\stochAction\ executed at $\state_i$ as if it has no effect at
states $\state$ if $\poss(\stochAction) \not\subseteq
\state$.

@@

An optimal solution to a finite-horizon POMDP is a contingent plan,
and can be expressed as a mapping from observation histories to
actions. Although suboptimal in general, useful plans can also take a
classical sequential format. This is the case in {\em conformant}
planning, where the objective is to find a sequence of actions that
achieves a goal ---i.e., reaches a state that satisfies some given
Boolean condition--- with probability $1$.  More generally, a useful
formalism for representing solutions to POMDPs is the finite-state
controller (FSC). This is a three-tuple $\langle \nodes, \selection,
\transition \rangle$ where: $\node \in \nodes$ is a set of controller
states, $\selection_\node(\action) = P(\action | \node)$ gives the
probability that the controller prescribes \action\ when in state
\node, and $\transition_\node(\action, \observation, \node') =
P(\node'|\node, \action, \observation)$ gives the probability the
controller transitions to state $\node'$, supposing we execute
\action\ in state \node\ receiving observation \observation. When
acting for finite $N$ steps from controller state $n$, the value of a
controller corresponds to the expected cumulative reward:
